Supplemental Materials, Methods and Results

### 2021-05-31 supp.Rmd
Ignore update, and small changes to Supp. Supp needs work now

```{r FigS1, fig.cap = "Examples of chameleon colour signal change in response to different social and predatory contexts"}

files <- list.files("./photos/ESM/")
img1 <- readJPEG(paste0("./photos/ESM/", files[1]))
img2 <- readJPEG(paste0("./photos/ESM/", files[2]))
img3 <- readJPEG(paste0("./photos/ESM/", files[3]))
img4 <- readJPEG(paste0("./photos/ESM/", files[4]))
img5 <- readJPEG(paste0("./photos/ESM/", files[5]))
img6 <- readJPEG(paste0("./photos/ESM/", files[6]))

p1 <- cowplot::ggdraw()+
        cowplot::draw_image(img1)
p2 <- cowplot::ggdraw()+
        cowplot::draw_image(img2)
p3 <- cowplot::ggdraw()+
        cowplot::draw_image(img3)
p4 <- cowplot::ggdraw()+
        cowplot::draw_image(img4)
p5 <- cowplot::ggdraw()+
        cowplot::draw_image(img5)
p6 <- cowplot::ggdraw()+
        cowplot::draw_image(img6)

(p1 + p3 + p6) / (p2 + p5 + p4) + plot_annotation(tag_levels = c("A", "B", "C", "D", "E", "F"))

@@

@@

## Supporting Supplementay Tables

```{r table1, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards in two social contexts (Male-Male Competition and Courtship). Standard errors are corrected by using robust variance estimators. "}

final_T1 <-  read.csv(file = "./output/tables/Table1_MS.csv")

 # Table
table1 <- flextable(final_T1) %>% table_style6()

table1
```

```{r table2, fig.width = 8, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates, standard errors (SE), Saitterwaite degrees of freedom (df) and 95% confidence internals (95% CI) comparing acromatic and chromatic mean differences between Kenyan and Hawaiian lizards for snakes and lizards. Full models are provided and standard errors are corrected by using robust variance estimators."}

 final_T2 <- read.csv(file = "./output/tables/Table2_MS.csv")
 # Table
table2 <- flextable(final_T2) %>%
           table_style7()

table2
```

```{r table3, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Planned contrasts between mean acromatric and chormatic contrast in Hawaii and Kenya for both predators (snakes and birds). Estimates of the mean difference ebtween Hawaii and Kenya propulations are provided along with their 95% confidence intervals (95% CI lower and upper). Wald tests of contrast significance are provided (F, degrees of freedom (df) and associated p-value).  We also applied a Bonferroni correction to p-value to control for multiple comparisons."}

contrast <- read.csv(file = "./output/tables/Table3_MS.csv")

table3 <- flextable(contrast) %>% table_style8()
table3
```

```{r table4, fig.width = 9, fig.height = 1, fig.asp = 0.1136364, tab.cap = "Model estimates (mean and contrasts), standard error (SE), degrees of freedom (df) and 95% confidence intervals (95% CI) for comparing acromatric and chromatic contrast of Hawaiian lizards against their introduced Hawaiian background (intercept) and Hawaiian lizards against their native Kenyan background (Background = Kenya) in both social contexts (Male-male competition and Courtship). Models were fit using robust variance estaimators to correct standard errors for non-independence, and we used a Saitterwaite degrees of freedom. Significance parameter estimates are those where 95% CI does not include zero."}

tb4 <- read.csv(file = "./output/tables/Table4_MS.csv")
table4 <- flextable(tb4) %>% table_style9()
table4
```

## Assessing changes in sexual signals across different body regions
We took measurements of spectral reflectance for different body regions of each chamaleon. In our primary analysis, we averaged across different body regions because we were interested only in whether visual signals overall were different between Kenya and Hawaii. This was justified given that, across the different body regions, effects were largely in a similar direction, and often of similar magnitude (Figures \@ref(fig:FigS2) - \@ref(fig:FigS4) and Tables \@ref(tab:tableS1)- \@ref(tab:tableS5)).

#### Body regions in social contexts (Male-male and Courtship Display) show different magnitude of responses, but similar directionality

Overall, different body regions explained significant variation in achromatic and chromatic contrast in both social contexts, whereas differences between Kenya and Hawaii only existed in achormatic contrast (Figure \@ref(fig:FigS2) & Table \@ref(tab:tableS1)). A significant interaction between body region and population was found for achormatic contrast when males were displaying to each other (Table \@ref(tab:tableS1)). However, in all cases Kenya showed a strong reduction in acromatic contrast relative to Hawaii lizards, despite some variation in the magnitude of effect (Figure \@ref(fig:FigS2)B & D; Table \@ref(tab:tableS2)).

```{r FigS2, fig.width = 13, fig.height = 12, fig.cap = "Chormatic (A, C) and Achromatic (B, D) contrast in just noticible differences (JND) for Hawaiian and Kenyan lizards relative to their environments. All four body regions (gular, mid-flank, tailbase and topflank) are shown averaged across stems and leaves found in their habitat. Mean JNDs are provided along with 95% confidence intervals. A & B represent male-male contest displays whereas C & D represent courtship displays"}

#Making a point plot with error bars split by body size and population 
#Compute raw means and standard errors for body part and population for dS and assigned this data into Display_plot_dS_dat
Display_plot_dS_dat <- Display_1 %>% #data for dS Chromatic constrast
  group_by(Population, BodyRegion) %>% 
  summarise( Mean = mean(dS, na.rm = T), 
               SE = sd(dS)/sqrt(length(unique(ID))-1),
            Upper = Mean + 1.96*SE,
            Lower = Mean - 1.96*SE)

Display_plot_dL_dat <- Display_1 %>% #data for dL Achromatic contrast
  group_by(Population, BodyRegion) %>% 
  summarise(Mean = mean(dL, na.rm = T), 
            SE = sd(dL)/sqrt(length(unique(ID))-1),
            Upper = Mean + 1.96*SE,
            Lower = Mean - 1.96*SE)

#Plotting Chromatic contrast (dS)
p1 <- ggplot(Display_plot_dS_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
  geom_point(size = 3, position = position_dodge(width=0.2)) +
  geom_line(position = position_dodge(width=0.2)) +
  geom_errorbar(aes(ymin = Lower, ymax = Upper), position = position_dodge(width=0.2), width = 0) + 
  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
  xlab("") + 
  ylab("Chromatic contrast (JND)") +
  theme_bw() +
  theme(legend.position = "none",
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    axis.text=element_text(size=18),
    axis.title=element_text(size=20,face="bold"),
    legend.text =element_text(size = 12),
    legend.title =element_text(size = 12, face = "bold"))

#Plotting Achromatic contrast (dL)
p2 <- ggplot(Display_plot_dL_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
  geom_point(size = 3, position = position_dodge(width=0.2)) +
  geom_line(position = position_dodge(width=0.2)) +
  geom_errorbar(aes(ymin = Lower, ymax = Upper), position = position_dodge(width=0.2), width = 0) + 
  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
  xlab("") + 
  ylab("Achromatic contrast (JND)") +
  theme_bw() +
  labs(color = "Body Region") + 
  theme(legend.position = c(0.85, 0.85),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    axis.text=element_text(size=18),
    axis.title=element_text(size=20,face="bold"),
    legend.text =element_text(size = 12),
    legend.title =element_text(size = 12, face = "bold"))

#Making a point plot with error bars split by body size and population 
#Compute raw means and standard errors for body part and population for dS and assigned this data into Display_plot_dS_dat
Courtship_plot_dS_dat <- Courtship %>% #data for dS Chromatic constrast
  group_by(Population, BodyRegion) %>% 
  summarise(Mean = mean(dS, na.rm = T), 
            SE = sd(dS)/sqrt(length(unique(ID))),
            Upper = Mean + 1.96*SE,
            Lower = Mean - 1.96*SE)

Courtship_plot_dL_dat <- Courtship %>% #data for dL Achromatic contrast
  group_by(Population, BodyRegion) %>% 
  summarise(Mean = mean(dL, na.rm = T), 
            SE = sd(dL)/sqrt(length(unique(ID))),
            Upper = Mean + 1.96*SE,
            Lower = Mean - 1.96*SE)

#Plotting Chromatic contrast (dS)
p3 <- ggplot(Courtship_plot_dS_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
  geom_point(size = 3, position = position_dodge(width=0.2)) +
  geom_line(position = position_dodge(width=0.2)) +
  geom_errorbar(aes(ymin = Lower, ymax = Upper), position = position_dodge(width=0.2), width = 0) + 
  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
  xlab("") + 
  ylab("Chromatic contrast (JND)") +
  theme_bw() +
  theme(legend.position = "none",
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    axis.text=element_text(size=18),
    axis.title=element_text(size=20,face="bold"),
    legend.text =element_text(size = 12),
    legend.title =element_text(size = 12, face = "bold"))

#Plotting Achromatic contrast (dL)
p4 <- ggplot(Courtship_plot_dL_dat, aes(x = Population, y = Mean, group = BodyRegion, color = BodyRegion)) +
  geom_point(size = 3, position = position_dodge(width=0.2)) +
  geom_line(position = position_dodge(width=0.2)) +
  geom_errorbar(aes(ymin = Lower, ymax = Upper), position = position_dodge(width=0.2), width = 0) + 
  scale_color_manual(values = wes_palette("Cavalcanti1")) + #Change color palette here see: https://github.com/karthik/wesanderson
  xlab("") + 
  ylab("Achromatic contrast (JND)") +
  theme_bw() +
  theme(legend.position = c(0.85, 0.85),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    axis.text=element_text(size=18),
    axis.title=element_text(size=20,face="bold"),
    legend.text =element_text(size = 12),
    legend.title =element_text(size = 12, face = "bold"))

p1 + p2 + p3 + p4 + plot_annotation(tag_levels = c('A', 'B', 'C', 'D')) & theme(plot.tag = element_text(face = 'bold', size=30))

```{r tableS1, tab.cap = "Wald F tests comparing overall significance of model parameters. Wald F tests were computed using robust variance estimators with a Saitterwaite correction to the degrees of freedom. Interactions were first tested and models refit to test the signifiacnce of main effects"}
fullTable <- rbind(wald_table_ConspicDisplayAC, 
                   wald_table_ConspicDisplayCC, 
                   wald_table_ConspicCourtshipAC,
                   wald_table_ConspicCourtshipCC)

fullTable <- fullTable %>% 
              mutate(Context = c("Male-Male Display", rep("", 7), "Courtship Display", rep("", 7)),
                     JND = c("Achromatic Contrast", rep("", 3), "Chromatic Contrast", rep("", 3), "Achromatic Contrast", rep("", 3), "Chromatic Contrast", rep("", 3)),
                     p_val = sapply(p_val, function(x) p_value(x))) %>% select(Context, JND, Variable, Fstat, df_num, df_denom, p_val)

S3 <- flextable(fullTable) %>%
        table_style2(.)
S3
```

```{r tableS2, tab.cap = "Linear mixed effects model results using robust variance estimators and Saitterwaite degrees of freedom correction.Social Context"}
robust_sc_table <- rbind(robust_ConspicDisplayAC,
                          robust_ConspicDisplayCC,
                          robust_ConspicCourtshipAC,
                          robust_ConspicCourtshipCC) %>% 
                    mutate(p_Satt = sapply(p_Satt, function(x) p_value(x)),
                           SocialContext = c("Male-Male Display", rep("", 17), 
                                             "Courtship Display", rep("", 17)),
                           JND = c("Achromatic Contrast", rep("", 8), 
                                   "Chromatic Contrast", rep("", 8), 
                                   "Achromatic Contrast", rep("", 8), 
                                   "Chromatic Contrast", rep("", 8)))

robust_sc_table$parameter <- rownames(robust_sc_table)

robust_sc_table <- robust_sc_table %>% select(SocialContext, JND, parameter, beta, SE, tstat, df, p_Satt)

tableS5 <- flextable(robust_sc_table) %>%  table_style4()
tableS5
```

>>>>

@@

# Supplmental Methods

### *Statistical Analysis*

Chromatic and achromatic contrast from visual models, measured in just noticable differences (JNDs), were analysed using linear mixed effects models in R using the `lme4` package (@Bateslme4) in combination with the package `clubSandwich` (@Pustejovsky21). For all individuals, we measured signals for multiple body regions, each of which were contrasted between multiple background measurements (i.e., top and bottoms of two leaves and two stems).Deriving multiple contrast for each individual body region and background resulted in a substantial number of comparisons, each of which were not completely independent of each other. To ensure that non-indpendence of data did not impact our inferences we used a Robust Variance Estimator (RVE) (using our final models to correct standard errors) and a Saitterwaite degrees of freedom correction. Across the different contexts (i.e., social contexts, predation and background environments), chromatic and acromatic contrasts were analysed in seperate models as response variables. For analyses involving social context, we included z-transformed snout-vent length (SVL), population of origin, and their interaction as fixed effects. For models involving predatory contexts and visual systems, we included the predator type (i.e, snake or bird), population of origin and their interaction as fixed effects. Finally, to test local adaptation, we used only chameleons from Hawaii and tested the extent to which chromatic and achromatic contrast varied depending on whether Hawaiian signals were obersved on a native Kenyan background or their current introduced habitat background in Hawaii. Relevant model interactions were first tested using Wald tests. If found to be non-significant (p > 0.1) we dropped the interaction and fit a main effects model. We then tested the significance of the main effect of 'Population', 'Predator' or 'Background' (depending on the model) using Wald tests with robust variance estimators and a Saitterwaite degrees of freedom correction.  For our main analyses we averaged estimates (marginalised) across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see Supplementary Results). In all cases, our results and conclusions were not impacted (see supplmentary code).

@@
